%% Template for EU deliverable, using the deliverable.sty style file

\documentclass[12pt,a4paper,twoside]{article}

%% common package
\usepackage[headers]{deliverable}
\usepackage{xspace}
\usepackage{verbatim}
\usepackage[usenames]{color}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{graphicx}
\usepackage{url}
\usepackage{array}
\usepackage{hyperref}
\usepackage{courier}
%%

%%insert here other packages needed by sections

%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Titlepage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% declaration of variables used in style
\deliverableDocnumber{D1.3}
\deliverableTitle{Software for dealing with compliant contacts}

\deliverableAuthor{Michael Mistry}
\deliverableResponsiblePartner{UB}
\deliverableAffiliation{% Insert here authors affiliations
 $^1$ UB
}

\deliverableReviewer{Name of internal reviewer}
\deliverableCoordinator{Francesco Nori}
\deliverableActivityNumber{n} %% n=1,..,10
\deliverableActivity{SW}
\deliverableDoctype{Deliverable} %% or Prototype
\deliverableClassification{Public} % or Consortium
\deliverableDistribution{Consortium} %
\deliverableStatus{Draft} % Draft or Final
\deliverableDeliveryDate{29/2/2016}
\deliverableFile{D1.2.pdf} % please do not use "-" in the name
\deliverableVersion{1.0}
\deliverableDate{Feb.~29, 2016}
\deliverableYear{2015}
\deliverablePages{\pageref{LastPage}}
\deliverableChangelog{v.1.0 & Feb 13, 2013 & First draft %%\\\hline
%%              v.2.0 & Feb 20, 2007 & Final version
}
\deliverableProjectStartingDate{1st March 2013}
\deliverableProjectEndDate{28th February 2017}
\deliverableProjectAcronym{CoDyCo}
\deliverableProjectTitle{Whole-Body Compliant Dynamical Contacts in Cognitive Humanoids}
\deliverableContractNumber{600716}
\deliverableProjectCoordinator{Istituto Italiano di Tecnologia}
\deliverableProjectUrl{www.codyco.eu}
\deliverableFrameworkProgramme{FP7}

\deliverableWorkpackage{deliv WP1}
\deliverableEditors{Michael Mistry}
\deliverableContributors{Morteza Azad, Elmar Rueckert}
\deliverableReviewers{}
\deliverableAbstract{This deliverable references the open-source software
  repositories, and corresponding documentation, developed under the CoDyCo
  project for dealing with compliant contacts.}
\deliverableReviewers{}

\deliverableKeywordList{CoDyCo, software, whole-body, compliant contacts,
  postural control, balancing, learning, contact parameter estimation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Sections
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% constants
\newcommand{\botegoCaps}{BOTEGO}
\newcommand{\certhCaps}{CERTH}
\newcommand{\cybionCaps}{CYBION}
\newcommand{\nuigCaps}{NUIG}
\newcommand{\ubitechCaps}{UBITECH}

%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% BEGIN DOCUMENT
\begin{document}

\deliverableMaketitle

%%TODO move to style
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

\textbf{Document Revision History}
\begin{center}
\begin{tabular}{|C{2cm}|C{3cm}|p{5cm}|C{4cm}|}
\hline
\textbf{Version}&\textbf{Date}&\textbf{Description}&\textbf{Author}\\\hline
First draft & 29/2/2016 & initial release & Michael Mistry\\\hline
\end{tabular}
\end{center}
 
 \clearpage

\newpage
\renewcommand*\contentsname{Table of Contents}
\renewcommand*\listfigurename{Index of Figures}
\tableofcontents
\newpage
\listoffigures
\newpage

%%%%%%%%%%%%%%%%%%%%%%%% Start deliverable content here.

\section{Software Modules}

All software developed for the CoDyCo project, including modules for the
control of balancing and reaching with multiple contacts, and also dealing
with compliant contacts are available as open-source within the following
repositories:
\begin{itemize}
\item \texttt{codyco-superbuild}: Main software repository.
\item \texttt{idyntree}: Dynamics library designed for free floating robots.
\item \texttt{wholebodyinterface}: C++ Interfaces to sensor measurements,
  state estimations, kinematic/dynamic model and actuators for a floating base
  robot.
\item \texttt{yarp-wholebodyinterface}: Implementation of the
  wholeBodyInterface for YARP robots.
\item \texttt{WB-Toolbox}: Simulink Toolbox for rapid prototyping of Whole
  Body Robot Controllers.
\item \texttt{codyco-modules}: Whole-body Compliant Dynamical Contacts in
  Cognitive Humanoids.
\item \texttt{mex-wholebodymodel}: Matlab MEX interface to the iWholeBodyModel
  interface.
\item \texttt{codyco-commons}: Miscellanous libraries developed in the CoDyCo
  project.
\item \texttt{WBI-Toolbox-controllers}: Controllers developed using the
  WB-Toolbox.
\item \texttt{LWR-ContactParams}: Locally Weighted Regression (LWR) algorithm
  for learning contact parameters of compliant contacts (this is available
  from \url{https://github.com/azadm/LWR_for_ContactParams.git}).
\end{itemize}

Further details and freely available software downloads are available via
github: \\ \url{https://github.com/robotology} \\

Additional documentation is available via the icub wiki: \\
\url{http://wiki.icub.org/codyco/dox/html/index.html} \\



\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% save-place: t
%%% End:
