%% Template for EU deliverable, using the deliverable.sty style file

\documentclass[12pt,a4paper,twoside]{report}

%% common package
\usepackage[headers]{deliverable}
\usepackage{xspace}
\usepackage{verbatim}
\usepackage[usenames]{color}
\usepackage[usenames,dvipsnames]{xcolor}
%\usepackage{graphicx}
\usepackage{url}
\usepackage{array}
\usepackage{graphics} 		% for pdf, bitmapped graphics files
\usepackage{graphicx}
\usepackage{epsfig} 		% for postscript graphics files
\usepackage{epstopdf}
\usepackage{caption}
\usepackage[labelformat=simple]{subcaption}
\renewcommand\thesubfigure{(\alph{subfigure})}
\usepackage{nameref}
\usepackage{multirow}
\usepackage{color}
%\usepackage{subfigure}
%%

%%insert here other packages needed by sections
%\usepackage{esint}
\usepackage{amstext}
\usepackage{amsmath}	 	% assumes amsmath package installed
\usepackage{amssymb}  		% assumes amsmath package installed
%\renewcommand\thesubfigure{(\alph{subfigure})}
\usepackage{wrapfig}
\usepackage[]{nomencl}		% nomenclatures
%\usepackage{dsfont}
%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Special Commands and Packages needed by Elmar
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{bm}
\usepackage{amsfonts}
\usepackage{setspace}
\usepackage{mathtools}
\newcommand{\FigureAbbr}{Fig~}%PLOS CB style
\newcommand{\FigureAbbrP}{Fig~}%PLOS CB style
\newcommand{\N}{\ensuremath{\mathcal{N}}}
\renewcommand{\vec}[1]{\ensuremath{\boldsymbol{#1}}}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Titlepage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% declaration of variables used in style
\deliverableDocnumber{D2.2}
\deliverableTitle{Models of human whole body motions in contact with rigid and compliant support surfaces}

\deliverableAuthor{Morteza Azad, Elmar Rueckert$^1$ and Jan Peters$^{1,2}$}
\deliverableResponsiblePartner{TUD}
\deliverableAffiliation{% Insert here authors affiliations
 $^1$ Intelligent Autonomous Systems Lab, Technische Universit\"at Darmstadt, 64289 Darmstadt, Germany.
  $^2$ Robot Learning Group, Max-Planck Institute for Intelligent Systems,
	Tuebingen, Germany.
}

\deliverableReviewer{}
\deliverableCoordinator{Jan Babi\v{c}}
\deliverableActivityNumber{n} %% n=1,..,10
\deliverableActivity{Activity Name}
\deliverableDoctype{Deliverable} %% or Prototype
\deliverableClassification{Public} % or Consortium
\deliverableDistribution{Consortium} %
\deliverableStatus{Draft} % Draft or Final
\deliverableDeliveryDate{29/2/2016}
\deliverableFile{Dxx\_DeliverableName.pdf} % please do not use "-" in the name
\deliverableVersion{1.0}
\deliverableDate{Feb.~29, 2016}
\deliverableYear{2016}
\deliverablePages{\pageref{LastPage}}
\deliverableChangelog{v.0.1 & Feb 04, 2016 & Initial draft %%\\\hline
%%              v.2.0 & Feb 20, 2007 & Final version
}
\deliverableProjectStartingDate{1st March 2013}
\deliverableProjectEndDate{28th February 2017}
\deliverableProjectAcronym{CoDyCo}
\deliverableProjectTitle{Whole-Body Compliant Dynamical Contacts in Cognitive Humanoids}
 \deliverableContractNumber{600716}
 \deliverableProjectCoordinator{Istituto Italiano di Tecnologia}
 \deliverableProjectUrl{www.codyco.eu}
 \deliverableFrameworkProgramme{FP7}
 
 \deliverableWorkpackage{WP2}
 \deliverableEditors{Morteza Azad and Jan Babi\v{c}}
 \deliverableContributors{Morteza Azad (UB) / Elmar Rueckert, Serena Invaldi, Jan Peters (TUDA) / Jernej Camernik, Jan Babic (JSI)}
 \deliverableReviewers{}
\deliverableAbstract{The scope of the current deliverable is to present the results on developing models of human whole body motions. 
}
\deliverableKeywordList{contacts, model learning, probabilistic movement representations}


\def\BibPath{./manuscript}       % location of bibtex-files
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Sections
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% constants
\newcommand{\botegoCaps}{BOTEGO}
\newcommand{\certhCaps}{CERTH}
\newcommand{\cybionCaps}{CYBION}
\newcommand{\nuigCaps}{NUIG}
\newcommand{\ubitechCaps}{UBITECH}

%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% BEGIN DOCUMENT
\begin{document}

\deliverableMaketitle

%%TODO move to style
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

\textbf{Document Revision History}
\begin{center}
\begin{tabular}{|C{2cm}|C{3cm}|p{5cm}|C{4cm}|}
\hline
\textbf{Version}&\textbf{Date}&\textbf{Description}&\textbf{Author}\\\hline
v. 0.1 & Feb. 04 & Initial Draft & Morteza Azad\\\hline
v. 1.0 & Feb. 29 & Final & Morteza Azad\\\hline
\end{tabular}
\end{center}
 
 \clearpage

\newpage
\renewcommand*\contentsname{Table of Contents}
\renewcommand*\listfigurename{Index of Figures}
\tableofcontents
\newpage
%\listoffigures
%\newpage

%%%%%%%%%%%%%%%%%%%%%%%% Start deliverable content here.

\chapter{Introduction}

%Short summary of our collaborative contribution
\paragraph{Postural Control Precedes and Predicts Volitional Motor Control (TUD/JSI)} 

The human motor control system is involved in most learning tasks including 
movement skill acquisition, cognitive reasoning and all flavors of communication 
skills like speech, sign language, gestures and writing. It processes perceptual 
streams, recalls memorized movements and is modulated by anticipated goals. To 
minimize the effect of all these influences, most studies investigate motor learning 
in unnatural conditions where the results are difficult to exploit. We show that 
motor adaptation can be analyzed under more ecological and thus more natural 
conditions by inspecting  the adaptation of a full-body motor task. A target had 
to be reached by the motion of one arm and the postural stability had to be 
maintained by the other arm. In a novel movement model we found strong  
correlations between both arms which were used to predict the reaching motion 
solely from observing the motion related to keeping postural stability. This 
finding has the potential to impact pre-tests of central nervous system 
disorders that are less prone to factors like stress, sleep deprivation and age 
compared to the classical cognitive tests. In robotics, the model can be 
exploited to overcome current limitations of autonomous robots in interacting 
with the environment through supportive contacts. 

\chapter{Executive Summary}

\chapter{Postural Control Precedes and Predicts Volitional Motor Control (TUD/JSI)}\label{sec:ElmarPrePrint}
\input{Elmar/PlosCBPreprint}

\bibliographystyle{plain}
\bibliography{manuscriptAll}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% save-place: t
%%% End:
