%!TEX root = ../../fourthYearReport.tex


 
\paragraph*{WP3: control and optimization of whole-body motion in contact (UPMC)}

After the fourth year of project, WP3 objectives were achieved for the fourth year. In summary, the main accomplishments and impacts for the research community are as follows: 

\begin{itemize}

\item[T3.3] IIT proposed a control laws ensuring the stabilization of a time-varying desired joint trajectory and joint limit avoidance in the case of fully-actuated manipulators.
\item[T3.3] As part of both WP1 and WP3, UPMC has pursued the generic implementation of an optimisation based whole-body controller named OCRA. 
\item[T3.4] TUD and INRIA, in collaboration with WP4, continued the collaboration on the topic of learning torque control in presence of multiple contacts, exploiting the force/torque and tactile sensors of iCub.
\item[T3.4] UB and TUD worked on developing a method to realize desired contact normal forces between a robot's end-effector and its compliant
environment.
\item[T3.4] INRIA designed a multi-task prioritized controller with soft task priorities.
\item[T3.4] UPMC developed a task compatibility optimization framework allowing for the incremental adaptation of operational tasks for whole-body control. This framework will be used for the implementation of year 4 demonstration.
\item[T3.4] UPMC pursued the generic implementation of a Model Predictive stepping approach for robust balance.
\item[WP2] JSI and UPMC kept developing a computational model of a reaching controller using the Speed-Accuracy trade-off.

 \end{itemize}

