%!TEX root = ../../fourthYearReport.tex


\paragraph{Work package 1 progress}

\subparagraph*{Software architecture design and evaluation of available
  open-source software pertinent to the scope of the project. (T1.1)}

The goal of T1.1 was to agree on a specific software architecture with
associated software tools whose specifications, dependencies and
interconnections meet the requirements and needs for achieving the goals of
the project.  The software, which is called \texttt{codyco-superbuild}, has
been available via github on
\texttt{https://github.com/robotology/codyco-superbuild} since the second year
of the project.  Details about the modules of the software are available in
deliverables D1.1, D1.2 and D1.3.

\subparagraph*{Simulator for whole-body motion with contacts (T1.2)}

The CoDyCo project requires a modular, component-based dynamics simulation
software providing numerically stable, computationally efficient and
physically consistent simulations of whole-body virtual human(oid) systems in
contact with rigid or soft environments.  During year four the iCub simulator
was further developed, keeping it aligned with the robot development.


\subparagraph{Control library for flexible specification of task space
  dynamics of floating base manipulators. (T1.3)} 

UPMC has kept developing OCRA over the 4th year of CoDyCo.  OCRA stands for
Optimization-based Control for Robotics Applications.  OCRA is a set of tools
which facilitates the development of optimization-based controllers for
robots.  At its core there is ocra-recipes, a group of platform independent
libraries which can be used to quickly develop optimization based controllers
for any robot.  Hierarchical, weighted, and hybrid controller schemes can
easily be implemented using the ocra-recipes libraries.  The generic
interfaces provided by OCRA allow different robots to use the exact same
controllers.  Examples of such implementations can be found for the humanoid
robot iCub (ocra-wbi-plugins), and the 7 DoF Kuka LWR (ocra-kdl).  OCRA also
allows users to specify high-level objectives via tasks.  These tasks provide
an intuitive way of generating complex behaviours and can be specified in XML
format.  In addition, a variety of gazebo plugins and controller visualization
tools have been developed to facilitate debugging and tuning of the
controller.

Among these plugins, a predictive approach \cite{ibanez2015Emergence} to
preview the duration and placement of coplanar contacts has been implemented
in the form of a client for OCRA using the iCub humanoid robot.  Within a
model-predictive control framework, the problem is formulated as a linearly
constrained mixed-integer quadratic program (MIQP) which allows the
determination over a preview horizon, of the optimal changes in the base of
support of the robot with compatible CoM behaviour, subject to multiple
constraints, while maximising balance and performance of a walking activity.



\subparagraph{System dynamics estimation software. Extension to
environmental compliance estimation (T1.4)}

The goal of T1.4 was to include compliant contact estimation to the software.
This is part of \texttt{codyco-superbuild} software which its details are
reported in deliverables D1.1, D1.2 and D1.3.


\subparagraph{Extension and enhancement of the iDyn library. (T1.5)}

As part of this task, INRIA developed several software modules which are all
available on github.  These modules which have been used as a support for the
research in the other WPs are:
\begin{itemize}
  \item software and GUI for visualizing the whole-body torques of iCub:
    \url{https://github.com/inria-larsen/icub-wholebody-visualization}
  \item software for interfacing the Geomagic software with Gazebo and use it
    for haptic interaction with iCub:
    \url{https://github.com/inria-larsen/geomagic_touch}
  \item software for learning prioritized controllers, in Matlab, with
    implementation of constrained stochastic optimization algorithms:
    \url{https://github.com/serena-ivaldi/learnOptimWBC}
  \item software for learning Probabilistic Movements Primitives from
    trajectories acquired by demonstrations on iCub:
    \url{https://github.com/inria-larsen/icubLearningTrajectories}
\end{itemize}
