%%%%%%%%%%%%%
%%%%JSI%%%%%%
%%%%%%%%%%%%%
\subsection{JSI contributions to dissemination}

2 invited talks, 1 editorial for journal special issue, 3 publications (2 journal, 1 internal conferences), 0 media coverage events.

\subsubsection{Invited talks}

\begin{enumerate}

\item Babic, Jan. Human-in-the-loop control of robots for industrial assembly tasks : invited talk, Omron Keihanna Technology Innovation Center, 7th September 2015, Kyoto, Japan.

\item Babic, Jan. Compliant robotic behaviour through human sensorimotor adaptation : presented at ICRA 2015, IEEE International Conference on Robotics and Automation, May 26th-30th, 2015, Seattle, Washington, USA.

\end{enumerate}

\subsubsection{Editorial work}

\begin{enumerate}
	
\item Babic, J. Guest editor of Autonomous robots, Kluwer Academic Publishers, 1994-. ISSN 0929-5593. http://link.springer.com/journal/10514.

\end{enumerate}

\subsubsection{Publications}

Ivaldi, S., Babic, J., Mistry, M., Murphy, R. Special issue on whole-body control of contacts and dynamics for humanoid robots. Autonomous robots, 2016, vol. 40, no. 3, 425-428.
 
Peternel, L., Noda, T., Petric, T., Ude, A,. Morimoto, J., Babic, J. Adaptive control of exoskeleton robots for periodic assistive behaviours based on EMG feedback minimisation. PloS one, ISSN 1932-6203, 2016, vol. 11, no. 2, 0148942-1-0148942-26.

Peternel, L., Petric, T., Babic, J. Human-in-the-loop approach for teaching robot assembly tasks using impedance control interface. In: 2015 IEEE International Conference on Robotics and Automation, May 26th-30th, 2015, Seattle, Washington, USA. ICRA 2015. Danvers: IEEE = Institute of Electrical and Electronics Engineers, cop. 2015, 1497-1502.