\subsection{UB contributions to dissemination}

1 invited talk, 1 editorial for journal special issue, 4 publications (1 journal, 2 international conferences, 1 workshop poster), 0 media coverage events. 

\subsubsection{Invited talks}

\begin{enumerate}
\item  Dr. Michael Mistry gave an invited workshop presentation on "Exploiting contact for whole-body physical human-robot collaboration" at the IEEE International Conference on Intelligent Robots and Systems (IROS 2015). 

\end{enumerate}

\subsubsection{International events participation}

\begin{enumerate}

\item  Morteza Azad and Michael Mistry presented the poster: Manipulability of the Center of Mass, at the Workshop on Interpersonal Postural Interactions, Birmingham, 22 June 2015.


\end{enumerate}

\subsubsection{Talks at international conferences}

\begin{enumerate}

\item  Morteza Azad presented the paper on "Balance control strategy for legged robots with compliant contacts" at ICRA 2015.

\item  Valerio Ortenzi presented the paper on "Projected inverse dynamics control and optimal control for robots in contact with the environment: A comparison" at IROS 2015


\end{enumerate}

\subsubsection{Editorial work}

\begin{enumerate}
	
\item Mistry, M. Guest editor of Autonomous robots, Kluwer Academic Publishers, 1994-. ISSN 0929-5593. http://link.springer.com/journal/10514.

\end{enumerate}


\subsubsection{Publications}

\begin{enumerate}
\item Ivaldi, S., Babic, J., Mistry, M., Murphy, R. Special issue on whole-body control of contacts and dynamics for humanoid robots. Autonomous robots, 2016, vol. 40, no. 3, 425-428.

\item Azad, M. and Mistry, M. Balance control strategy for legged robots with compliant contacts.  IEEE International Conference on Robotics and Automation (2015), 4391-4396.

\item Ortenzi V, Stolkin R, Kuo JA, Mistry M. Projected inverse dynamics control and optimal control for robots in contact with the environment: A comparison, IEEE/RSJ International Conference on Intelligent Robots and Systems (2015), 4009-4015.

\item Azad, M. and Mistry, M. Manipulability of the Center of Mass, Poster at the Workshop on Interpersonal Postural Interactions, Birmingham, 22 June 2015.

\end{enumerate}
	

