\subsection{UPMC contributions to dissemination}

7 invited talks, 4 other events, 5 paper presentations at international Conferences, 1 international journal, 5 international conferences.

\subsubsection{Invited talks}

\begin{enumerate}

\item  V. Padois. Robotique industrielle: évolution, enjeux et perspectives, January 2016. Invited talk at CNER/SERECT.

\item N. Perrin. Quelles mathématiques pour le contrôle du mouvement humanoïde ?, January 2016. Invited Seminar at the Math-club of University Paris 7.

\item O. Sigaud. Des robots immergés dans la société ? November 2015. Invited talk at the CE Industriel d’Air France.

\item V. Padois. Design and control of collaborative robots: a focus on ergonomics and safety, November 2015. Panel Session on ``360◦ on CNRS vision of FoF and Smart Cities'' at the Colloque From Industry4.0 to Smart Cities.

\item V. Padois. Design and control of collaborative robots: a focus on ergonomics and safety, October 2015. Invited talk at the National Days on Robotics Research.

\item V. Padois. Whole-body compliant dynamical contacts for humanoids: the codyco project, June 2015. Invited talk at the National Days on Humanoid Robotics.

\item O. Sigaud.  Des robots immergés dans la société ? Mars, 2015. Invited seminar at la Semaine du Cerveau à la Faculté Saint-Charles de Marseille.
\end{enumerate}

\subsubsection{Other events}

\begin{enumerate}

\item  Darwin Lau. Generation of Dynamically Balanced Locomotion with Multiple Non-coplanar Contacts, June 2015. Presented at the National Days on Humanoid Robotics.

\item Ryan Lober. Tasks compatibility in whole-body control, June 2015. Presented at the National Days on Humanoid Robotics.

\item Aurélien Ibanez defended his PhD thesis on "Emergence of complex behaviors from coordinated predictive control in humanoid robotics" in september 2015.

\item Découverte du robot humanoïde iCub, October 2015. French national days for Science.

\end{enumerate}

\subsubsection{Talks at international conferences}

\begin{enumerate}
\item  Mingxing Liu presented the paper on "Generalized Projector for Task Priority Transitions During Hierarchical Control" at ICRA 2015.

\item Darwin Lau presented the paper on "Effective Generation of Dynamically Balanced Locomotion with Multiple Non-coplanar Contacts" at ISRR 2015.

\item Vincent Padois presented the paper on "Reactive whole-body control for humanoid balancing on non-rigid unilateral contacts" at IROS 2015.

\item Ryan Lober presented the paper on "Variance Modulated Task Prioritization in Whole-Body Control" at IROS 2015.

\item Darwin Lau presented the paper on "Minimization of the rate of change in torques during motion and force control under discontinuous constraints" at Robio 2015.
\end{enumerate}


\subsubsection{Publications}

\begin{enumerate}
\item M. Liu, S. Hak, and V. Padois. Generalized projector for task priority transitions during hierarchical control. In Proceedings of the IEEE International Conference on Robotics and Automation, pages 768–773, Seattle, USA, May 2015.

\item M. Liu, R. Lober, and V. Padois. Whole-body hierarchical motion and force control for humanoid robots. Autonomous Robots, 40(3):493–504, 2016.

\item M. Liu and V. Padois. Reactive whole-body control for humanoid balancing on non-rigid unilateral contacts. In Proceedings of the IEEE/RSJ International Conference on Intelligent Robots and Systems, pages 3981–3987, Hamburg, Germany, September 2015.

\item R. Lober, V. Padois, and O. Sigaud. Variance modulated task prioritization in whole-body control. In Proceedings of the IEEE/RSJ International Conference on Intelligent Robots and Systems, pages 3944–3949, Hamburg, Germany, September 2015.

\item N. Perrin, D. Lau, and V. Padois. Effective generation of dynamically balanced locomotion with multiple non-coplanar contacts. In Proceedings of the International Symposium on Robotics Research, Sestri Levante, Italy,
September 2015.

\item Y. Tan, D. Lau, M. Liu, P. Bidaud, and V. Padois. Minimization of the rate of change in torques during motion and force control under discontinuous constraints. In Proceedings of the IEEE International Conference on Robotics and Biomimetics (Robio), pages 2621–2628, Zhuhai, China, December 2015.
\end{enumerate}
	
