%!TEX root = ../../secondYearReport.tex


\paragraph{Work package 7 progress}

Dissemination and exploitation activities included the participation to international events addressed to both commercial and academic institutions. 

\subparagraph{Dissemination activities towards academia, industry, and other users (T7.1)}

Dissemination activities were conducted thorough international publications, organisation of international events, talks at international conferences, press interviews and iCub expositions at international events. Here is the overall contribution subdivided by partner:

\begin{itemize}

\item IIT: 4 invited talks, 4 organised international events, 6 talks at international conferences, 10 publications (2 journal, 7 internal conferences, 1 book chapter), 10 media coverage events.

\item TUD: 15 invited talks, 4 organised international events, 7 publications (7 internal conferences), 7 media coverage events.

\item UPMC: 3 invited talks, 6 publications (6 internal conferences), 1 media coverage event.

\item UB: 3 invited talks, 5 publications (5 internal conferences), 5 talks at international conferences.

\item JSI: 1 invited talks, 1 organised special issue, 2 talks at international conferences, 7 publications (1 journal, 6 internal conferences).

\item INRIA: 4 invited talks, 4 organised international events, 8 publications (4 journal, 4 internal conferences), 3 media coverage events

\end{itemize}

\subparagraph{Exploitation plan (T7.2)}

The second year activities on T7.1 and T7.2 are all contained in ``D7.1 Dissemination and exploitation plan'' available here: \url{https://github.com/robotology-playground/codyco-deliverables/tree/master/D7.1/pdf}.

\subparagraph{Management of IPR (T7.3)}

No activities to be reported during the second year on this task in consideration of the fact that the task started at the very end of the second year. As a minor starting activity the consortium circulated a list containing each partner responsible contact person for the IPR management. This list is contained in ``D7.1 Dissemination and exploitation plan'' available here: \url{https://github.com/robotology-playground/codyco-deliverables/tree/master/D7.1/pdf}.

\subparagraph{Dissemination of a database of human motion with contacts (T7.4)}

During the second year of CoDyCo, IIT completed the task of setting up a database for storing both human and robot datasets. The details on the database are reported in ``D7.2 Standard database with support materials'' available here \url{https://github.com/robotology-playground/codyco-deliverables/tree/master/D7.2/pdf}. 

