%% Template for EU deliverable, using the deliverable.sty style file

\documentclass[12pt,a4paper,twoside]{article}

%% common package
\usepackage[headers]{deliverable}
\usepackage{xspace}
\usepackage{verbatim}
\usepackage[usenames]{color}
\usepackage{listings}
\usepackage[usenames,dvipsnames,table]{xcolor}
\usepackage[pdftex,dvips]{graphicx}
\usepackage{url}
\usepackage{array}

%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Misc. packages
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\graphicspath{{./figure/}}
\usepackage{pdfpages}
\usepackage{amsmath} % assumes amsmath package installed
\usepackage{amssymb}  % assumes amsmath package installed
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{hyperref}
\hypersetup{
	bookmarks=true,         % show bookmarks bar?
	unicode=false,          % non-Latin characters in Acrobat’s bookmarks
	pdftoolbar=true,        % show Acrobat’s toolbar?
	pdfmenubar=true,        % show Acrobat’s menu?
	pdffitwindow=false,     % window fit to page when opened
	pdfstartview={FitH},    % fits the width of the page to the window
	pdfkeywords= {}, % list of keywords
	pdfnewwindow=true,      % links in new window
	colorlinks=true,       % false: boxed links; true: colored links
	linkcolor=black,          % color of internal links (change box color with linkbordercolor)
	citecolor=black,        % color of links to bibliography
	filecolor=black,      % color of file links
	urlcolor=black           % color of external links
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Luka's packages
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{cite}
\usepackage{mathptmx} % assumes new font selection scheme installed
\usepackage{bm,times} % assumes new font selection scheme installed
\usepackage{amsfonts}
\usepackage{stfloats}
\usepackage{psfrag}
\usepackage{float}
\usepackage{color}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Serena's packages
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\pdfoutput=1

\RequirePackage{fix-cm}
\usepackage{multicol}
\usepackage{xcolor,colortbl}
\usepackage{wrapfig}
\usepackage{mdwtab} %added by Na
\usepackage{mdwmath}
\usepackage{enumerate}
\usepackage{algorithm2e}
\usepackage[utf8]{inputenc}
\usepackage{caption}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Titlepage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% declaration of variables used in style
\deliverableDocnumber{D2.3}
\deliverableTitle{Models of systematic application of physical interaction with the environment and humans}

\deliverableAuthor{Jernej \v{C}amernik$^{1}$, and Jan Babi\v{c}$^{1}$}
\deliverableResponsiblePartner{JSI}
\deliverableAffiliation{$^1$ Department for Automation, Biocybernetics and Robotics, Jo\v{z}ef
  Stefan Institute, Ljubljana, Slovenia.}

\deliverableReviewer{}
\deliverableCoordinator{Jan Babi\v{c}}
\deliverableActivityNumber{n} %% n=1,..,10
\deliverableActivity{Activity Name}
\deliverableDoctype{Deliverable} %% or Prototype
\deliverableClassification{Public} % or Consortium
\deliverableDistribution{Consortium} %
\deliverableStatus{Final} % Draft or Final
\deliverableDeliveryDate{28/2/2017}
\deliverableFile{D2.3\D2.3.pdf} % please do not use "-" in the name
\deliverableVersion{1.0}
\deliverableDate{28/2/2017}
\deliverableYear{2017}
\deliverablePages{\pageref{LastPage}}
\deliverableChangelog{v.0.1 & Feb 15, 2017 & Initial draft %%\\\hline
%%              v.2.0 & Feb 20, 2007 & Final version
}
\deliverableProjectStartingDate{1st March 2013}
\deliverableProjectEndDate{28th February 2017}
\deliverableProjectAcronym{CoDyCo}
\deliverableProjectTitle{Whole-Body Compliant Dynamical Contacts in Cognitive Humanoids}
\deliverableContractNumber{600716}
\deliverableProjectCoordinator{Istituto Italiano di Tecnologia}
\deliverableProjectUrl{www.codyco.eu}
\deliverableFrameworkProgramme{FP7}

\deliverableWorkpackage{WP2}
\deliverableEditors{Jernej \v{C}amernik, Michael Mistry and Jan Babi\v{c}}
\deliverableContributors{Serena Ivaldi (INRIA), Chie Takahashi (UB), Zrinka Potocanac (JSI), Luka Peternel (JSI), Jan Babi\v{c} (JSI)}
\deliverableReviewers{}
\deliverableAbstract{The scope of the current deliverable is to present the
  results on developing models of human whole body motions in contact with
  environment.}
\deliverableKeywordList{contacts, model learning, probabilistic movement
  representations}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Sections
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% constants
\newcommand{\botegoCaps}{BOTEGO}
\newcommand{\certhCaps}{CERTH}
\newcommand{\cybionCaps}{CYBION}
\newcommand{\nuigCaps}{NUIG}
\newcommand{\ubitechCaps}{UBITECH}

% Chie
\newcommand{\BF}{\mathbf{F}}
\newcommand{\BZ}{\mathbf{Z}}
\newcommand{\BA}{\mathbf{A}}
\newcommand{\BB}{\mathbf{B}}
\newcommand{\BC}{\mathbf{C}}
\newcommand{\BD}{\mathbf{D}}
\newcommand{\BW}{\mathbf{W}}
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% BEGIN DOCUMENT
\begin{document}

\deliverableMaketitle

%%TODO move to style
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

\textbf{Document Revision History}
\begin{center}
\begin{tabular}{|C{2cm}|C{3cm}|p{5cm}|C{4cm}|}
\hline
\textbf{Version}&\textbf{Date}&\textbf{Description}&\textbf{Author}\\\hline
v. 1.0 & 28 Feb 2017 & Final Version & Jernej \v{C}amernik\\\hline
\end{tabular}
\end{center}
 
 \clearpage

\newpage
\renewcommand*\contentsname{Table of Contents}
%\renewcommand*\listfigurename{Index of Figures}
\tableofcontents
\newpage
%\listoffigures
%\newpage

%%%%%%%%%%%%%%%%%%%%%%%% Start deliverable content here.

\section{Introduction}

This deliverable reports on adaptive models describing changes in human whole body motion in contact with unpredictable support surfaces and humans. The goal of the work reported in this deliverable was to understand the factors involved in human choice of contact utilization and to investigate how interaction through contacts can contribute to learning of whole body motor control. There are numerous situations where humans utilize physical interaction with either the environment or another human to stabilise one’s own balance. These can be through light touch to gain additional haptic information from the environment; through weight-bearing support; through energy storage and recovery from a springy surface; and through assistive forces from other humans. In this regard we investigated how human physical interactions with rigid, compliant or spring-like contacts contribute to changes in whole body motion and to task-relevant parameters. The core question that we addressed is how a human can learn to predict the environmental properties by performing tasks with hand or body contacts and then generalize the learned predictive mechanisms to other tasks. For example, does interaction with the support surface during a supported reaching action allow effective choice of and use of the support? The findings reported in this deliverable are models that can be further used to enhance robot-learning capabilities by systematic application of physical interaction with the environment or humans.



% In the section tapata we.....
The major role of this work is to explore the mechanism which helps humans to learn the compliant force dynamics. This would enable us to get an in-depth understanding of how humans modulate their whole-body motion to reach anticipated goals ruled by the external force. Further, this would be beneficial to apply it for controlling the humanoid robot in order to keep a stable posture when interacting with multiple compliant surfaces. With reference to the results of the pilot experiment (see the D2.2), the experimental design has been developed and further human subject experiments were conducted to examine the human learnability factor in different force dynamics. Three different types of forces (simple linear and two non-linear forces) were generated and the human movements to reach the anticipated goal were measured against the compliant forces.

% In the section tapata we.....
Estimating the engagement is critical for human - robot interaction. Engagement measures typically rely on the dynamics of the social signals exchanged by the partners, especially speech and gaze. However, the dynamics of these signals is likely to be influenced by individual and social factors, such as personality traits, as it is well documented that they critically influence how two humans interact with each other. Here, we assess the influence of two factors, namely extroversion and negative attitude toward robots, on speech and gaze during a cooperative task, where a human must physically manipulate a robot to assemble an object. We evaluate if the score of extroversion and negative attitude towards robots co-variate with the duration and frequency of gaze and speech cues. The experiments were carried out with the humanoid robot iCub and N=56 adult participants. We found that the more people are extrovert, the more and longer they tend to talk with the robot; and the more people have a negative attitude towards robots, the less they will look at the robot face and the more they will look at the robot hands where the assembly and the contacts occur.
Our results confirm and provide evidence that the engagement models classically used in human-robot interaction should take into account attitudes and personality traits.

% In the section tapata we.....
We propose a novel method that arbitrates the control between the human and the robot actors in a teaching-by-demonstration setting to form synergy between the two and facilitate effective skill synthesis on the robot. We employed the human-in-the-loop teaching paradigm to teleoperate and demonstrate a complex task execution to the robot in real-time. As the human guides the robot to perform the task, the robot obtains the skill online during the demonstration. To encode the robotic skill we employed Locally Weighted Regression that fits local models to specific state region of the task based on the human demonstration. If the robot is in the state region where no local models exist, the control over the robotic mechanism is given to the human to perform the teaching. When local models are gradually obtained in that region, the control is given to the robot so that the human can examine its performance already during the demonstration stage, and take actions accordingly. This enables a co-adaptation between the agents and contributes to a faster and more efficient teaching. As a proof-of-concept, we realised the proposed robot teaching system on a haptic robot with the task of generation of a desired vertical force on a horizontal plane with unknown stiffness properties.

% In the section tapata we.....
Summary of what Zrinka did. If there is an additional uncertainty in balance (for example by perturbing balance with an underlying Stewart platform), how does contact preference change? 
These models will be further used to help the robot controller when the robot has joint torque limits, joint position limits, or narrow stability margins. In effect, the models will help the robot decide when to make or break the contacts.



%Short summary of our collaborative contribution

% \chapter{Executive Summary}
\newpage{}
\section{Compliant-contact force perception study}\label{sec:Chie}
\setcounter{figure}{0}
\input{Chie/ChieD2.3}
\clearpage{}

\section{Towards engagement models that consider individual factors in HRI: on the relation of extroversion and negative attitude towards robot to gaze and speech during a human-robot assembly task}\label{sec:Serena}
\setcounter{figure}{0}
\input{Serena/IJSR_edhhi}
\clearpage{}

\section{A Shared Control Method for Online Human-in-the-Loop Robot Learning Based on Locally Weighted Regression}\label{sec:Luka}
\setcounter{figure}{0}
\input{Luka/IROS2016}
\clearpage

\section{Stepping in a novel balance environment}\label{sec:Zrinka}
\setcounter{figure}{0}
\input{Zrinka/platform}
\clearpage{}

\bibliographystyle{plain}
\bibliography{platform,IROS2016,edhhi,ChieBib}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% save-place: t
%%% End:
