%% Template for EU deliverable, using the deliverable.sty style file

\documentclass[12pt,a4paper,twoside]{report}

%% common package
\usepackage[headers]{deliverable}
\usepackage{xspace}
\usepackage{verbatim}
\usepackage[usenames]{color}
\usepackage[usenames,dvipsnames]{xcolor}
%\usepackage{graphicx}
\usepackage{url}
\usepackage{array}
\usepackage{graphics} 		% for pdf, bitmapped graphics files
\usepackage{graphicx}
\usepackage{epsfig} 		% for postscript graphics files
\usepackage{epstopdf}
\usepackage{caption}
\usepackage[labelformat=simple]{subcaption}
\renewcommand\thesubfigure{(\alph{subfigure})}
\usepackage{nameref}
\usepackage{multirow}
\usepackage{color}
%\usepackage{subfigure}
%%

%%insert here other packages needed by sections
%\usepackage{esint}
\usepackage{amstext}
\usepackage{amsmath}	 	% assumes amsmath package installed
\usepackage{amssymb}  		% assumes amsmath package installed
%\renewcommand\thesubfigure{(\alph{subfigure})}
\usepackage{wrapfig}
\usepackage[]{nomencl}		% nomenclatures
%\usepackage{dsfont}
%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Special Commands and Packages needed by Elmar
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{bm}
\usepackage{amsfonts}
\usepackage{setspace}
\usepackage{mathtools}
\newcommand{\FigureAbbr}{Fig~}%PLOS CB style
\newcommand{\FigureAbbrP}{Fig~}%PLOS CB style
\newcommand{\N}{\ensuremath{\mathcal{N}}}
\renewcommand{\vec}[1]{\ensuremath{\boldsymbol{#1}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Titlepage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% declaration of variables used in style
\deliverableDocnumber{D2.2}
\deliverableTitle{Models of human whole body motions in contact with rigid and
  compliant support surfaces}

\deliverableAuthor{Morteza Azad$^{1}$, Michael Mistry$^{1}$ and Jan
  Babi\v{c}$^{2}$}
\deliverableResponsiblePartner{JSI}
\deliverableAffiliation{$^1$ School of Computer Science, University of
  Birmingham, UK.
  $^2$ Department for Automation, Biocybernetics and Robotics, Jo\v{z}ef
  Stefan Institute, Ljubljana, Slovenia.}

\deliverableReviewer{}
\deliverableCoordinator{Jan Babi\v{c}}
\deliverableActivityNumber{n} %% n=1,..,10
\deliverableActivity{Activity Name}
\deliverableDoctype{Deliverable} %% or Prototype
\deliverableClassification{Public} % or Consortium
\deliverableDistribution{Consortium} %
\deliverableStatus{Final} % Draft or Final
\deliverableDeliveryDate{29/2/2016}
\deliverableFile{Dxx\_DeliverableName.pdf} % please do not use "-" in the name
\deliverableVersion{1.0}
\deliverableDate{29/2/2016}
\deliverableYear{2016}
\deliverablePages{\pageref{LastPage}}
\deliverableChangelog{v.0.1 & Feb 04, 2016 & Initial draft %%\\\hline
%%              v.2.0 & Feb 20, 2007 & Final version
}
\deliverableProjectStartingDate{1st March 2013}
\deliverableProjectEndDate{28th February 2017}
\deliverableProjectAcronym{CoDyCo}
\deliverableProjectTitle{Whole-Body Compliant Dynamical Contacts in Cognitive Humanoids}
\deliverableContractNumber{600716}
\deliverableProjectCoordinator{Istituto Italiano di Tecnologia}
\deliverableProjectUrl{www.codyco.eu}
\deliverableFrameworkProgramme{FP7}

\deliverableWorkpackage{WP2}
\deliverableEditors{Morteza Azad, Michael Mistry and Jan Babi\v{c}}
\deliverableContributors{Morteza Azad (UB), Michael Mistry (UB), Chie
  Takahashi (UB), Elmar Rueckert (TUD), Jan Peters (TUD), Jernej \v{C}amernik
  (JSI), Jan Babi\v{c} (JSI)}
\deliverableReviewers{}
\deliverableAbstract{The scope of the current deliverable is to present the
  results on developing models of human whole body motions in contact with
  environment.}
\deliverableKeywordList{contacts, model learning, probabilistic movement
  representations}


\def\BibPath{./manuscript}       % location of bibtex-files
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Sections
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% constants
\newcommand{\botegoCaps}{BOTEGO}
\newcommand{\certhCaps}{CERTH}
\newcommand{\cybionCaps}{CYBION}
\newcommand{\nuigCaps}{NUIG}
\newcommand{\ubitechCaps}{UBITECH}

%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% BEGIN DOCUMENT
\begin{document}

\deliverableMaketitle

%%TODO move to style
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

\textbf{Document Revision History}
\begin{center}
\begin{tabular}{|C{2cm}|C{3cm}|p{5cm}|C{4cm}|}
\hline
\textbf{Version}&\textbf{Date}&\textbf{Description}&\textbf{Author}\\\hline
v. 1.0 & 29 Feb 2016 & Final Version & Morteza Azad\\\hline
\end{tabular}
\end{center}
 
 \clearpage

\newpage
\renewcommand*\contentsname{Table of Contents}
\renewcommand*\listfigurename{Index of Figures}
\tableofcontents
\newpage
%\listoffigures
%\newpage

%%%%%%%%%%%%%%%%%%%%%%%% Start deliverable content here.

\chapter{Introduction}

%Short summary of our collaborative contribution

This deliverable summarizes the contribution of the CoDyCo consortium in tasks
T2.2 and T2.3 at the end of the third year.  These tasks are \textit{design of
  models for human whole body motion in contact}, and \textit{strategies of
  dealing with uncertainties in contact}, respectively.  The results are
briefly explained in four chapters.  In chapter 2, dynamic manipulability of
the centre of mass (CoM) is introduced as a metric for measuring the balance
ability of legged robots while they are in contact with their environment.
Experiments on human subjects show the applicability of this theory in
analysing balancing recovery motion in humans.  In chapter 3, the effects of
using handles is studied for posture control of standing subjects while they
are perturbed by external forces.  It is shown that the use of handles
significantly reduces the displacement of the centre of pressure.  Also, it is
observed that subjects clearly relied on using the handle for support, even
though the perturbations did not pose a significant balance threat.  Chapter 4
investigates the use of supporting contact (and its relationship with task)
while it is required for accomplishing a task.  In the experiments, human
subjects are asked to reach a target by the motion of one arm while they has
to maintain their postural balance with the other arm.  In a novel movement
model, strong correlations between both arms are found which were used to
predict the reaching motion solely from observing the motion related to
keeping postural stability. This finding has the potential to impact pre-tests
of central nervous system disorders that are less prone to factors like
stress, sleep deprivation and age compared to the classical cognitive tests.
In robotics, the model can be exploited to overcome current limitations of
autonomous robots in interacting with the environment through supportive
contacts.

% \chapter{Executive Summary}

\chapter{Manipulability of the Center of Mass: A Tool to Study, Analyse and
  Measure the Ability to Balance (UB/JSI)}\label{sec:Morteza}
\input{Morteza/manipulability}

\chapter{Postural Control Precedes and Predicts Volitional Motor Control
  (TUD/JSI)}\label{sec:ElmarPrePrint}
\input{Elmar/PlosCBPreprint}

\chapter{Use of rigid contacts during continuous perturbations
  (JSI)}\label{sec:Jernej}
\input{Jernej/rigidContacts}

\bibliographystyle{plain}
\bibliography{manuscriptAll,rigidContacts}




\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% save-place: t
%%% End:
