%!TEX root = ../../secondYearReport.tex


\paragraph*{WP1: toolbox for computing and controlling dynamics of whole-body movements with contacts (UB)}

WP1 objectives were achieved for the second year. In summary, the main accomplishments and impacts for the research community are as follows: 

\begin{itemize}

\item A survey of over 100 robotics researchers worldwide was conducted on use of simulator tools.  To date, there has been no such objective evaluation of the various simulation packages available for conducting robotics research. The output of such a survey will help researchers determine the best tool for their application and thus, is of great value to the community.  A report of survey results was released as part of deliverable D1.1 and was submitted for journal publication. 

\item A new iCub simulator to cope with multi-contact dynamics has been released. The new simulator can now interface with Gazebo, a widely used and actively developed simulation engine. The simulator has been documented as part of deliverable D1.1

\item A tool for automatic generation of URDF models of digital humans has been developed. As URDF is a standardized format for representing articulated rigid bodies, researchers  now have ease to create human kinematic and dynamic models for simulation and analysis. Documentation of the tool was released as part of D1.1. 

\item New methods for whole-body dynamics estimation have been developed, and submitted for publication. 

\item A new Gaussian process (GP) model was developed that is explicitly designed to deal with non-linearities induced through contacts with the environment. This work was submitted for publication. 
 
 \end{itemize}

