%!TEX root = ../../thirdYearReport.tex


\paragraph*{WP4: adaptation, generalization and improvement of compliant control and tasks with contacts (TUD)}

% The goal of WP4 is to endow the CoDyCo humanoid robot control architecture with
% the core abilities for the adaptation, generalization and self-improvement of
% both control laws and tasks that involve physical interaction with humans, and
% the environment.

In T4.2 TUD and JSI studied the effect of supportive contacts on reaching 
motions. Through studies in probabilistic models we could show that these 
supportive contacts predict volitional movements. In ongoing work, we 
investigate if the derived concepts can be used for robot control. A pre-print 
was attached to D2.2 and the the paper is in revision at Scientific Reports 
(Nature Publishing Group). In another, more theoretical study, TUD showed for 
the first time that spiking neural networks can implement motion planning in 
continuous spaces. This research was published with Scientific Reports 
\cite{Rueckert_SR_2016}. 

In T4.4 TUD developed a model-free probabilistic movement primitive framework to 
learn variable stiffness controller from demonstrations of joint angle 
trajectories and joint forces. This work was tested in the iCub robot and was 
presented at an international robotics conference \cite{paraschos2015model}. 
UPMC continued their research on whole-body controller for humanoid robots. A 
novel control strategy for concurrently executing multiple tasks was developed. 
The controller can overcome incompatibilities between the tasks due to 
conflicting task objectives and was presented at an international robotics 
conference \cite{lober2015variance}.  